---
title: 'Sad Plot #1 Star Wars Baby Names'
author: Alex Michel
date: '2018-06-18'
slug: sad-plot-1-star-wars-baby-names
categories:
  - blog
tags:
  - R Markdown
  - sad plot
  - Star Wars
  - baby names
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

### 1) Clean Reproducible Code

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(babynames)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(plotly)
library(RColorBrewer)
```

```{r}
star_wars <- babynames %>%
  filter(name == "Leah" | name == "Leia") %>% 
  filter(sex == "F") %>% 
  filter(year >= 1975)

plot <- ggplot(star_wars, aes(x = year, 
                           y = prop,  
                           group = name,
                           color = name)) + 
  geom_line(size = 2) +
  geom_vline(xintercept = 1977, linetype="dashed", 
                color = "blue", size=1) +
  geom_vline(xintercept = 1999, linetype="dashed", 
                color = "blue", size=1) +
  geom_vline(xintercept = 2014, linetype="dashed", 
                color = "blue", size=1) +
  annotate("text", x = 1980, y = 0.01, label = "Star Wars") +
  annotate("text", x = 2004, y = 0.008, label = "Phantom Menace") +
  annotate("text", x = 2009.5, y = 0.005, label = "Force Awakens")

plot


```

### 2) What Type of Graph?

This is a line graph. 


### 3) Data Description

This graph was pretty much my first time graphing something in ggplot2! 

The data came from the babynames package. 

### 4) Intended Audience

Honestly there isn't much of an intended audience for this one (me?). I think it could be most useful as an accompanying illustration for an article on the rising popularity of Star Wars baby names.

### 5) What Am I Trying to Show?

1) Do Star Wars baby names become more popular after a movie comes out? (not really)


### 6) How Do I Read This?

The main components of the graph are time and proportion of baby names. As the proportion increases the line also increases. The use of time allows us to see whether the change in proportion is associated with particular points in time.

### 7) Presentation Tips

I would definitely include more fun images in this.


### 8) Variations and Alternatives

An alternative to this line graph could be a bar graph -- which would make it easier to see changes associated with particular years.


### 9) How I Created It: Methods Section

This was pretty much the first time I've ever used ggplot2! I started off by copying Alison's example and filtering the data for some simple name variations (Leah vs Leia) on spelling. I filtered by sex and year (only including years after Star Wars came out).


```{r}
star_wars <- babynames %>%
  filter(name == "Leah" | name == "Leia") %>% 
  filter(sex == "F") %>% 
  filter(year >= 1975)
```

My plot is pretty sad, but I included some annotations to show when various movies came out so that the viewer can figure out if there is a correlation.

```{r}
plot <- ggplot(star_wars, aes(x = year, 
                           y = prop,  
                           group = name, 
                           color = name)) + 
  geom_line(size = 2) +
  geom_vline(xintercept = 1977, linetype="dashed", 
                color = "blue", size=1) +
  geom_vline(xintercept = 1999, linetype="dashed", 
                color = "blue", size=1) +
  geom_vline(xintercept = 2014, linetype="dashed", 
                color = "blue", size=1) +
  annotate("text", x = 1980, y = 0.01, label = "Star Wars") +
  annotate("text", x = 2004, y = 0.008, label = "Phantom Menace") +
  annotate("text", x = 2009.5, y = 0.005, label = "Force Awakens")

plot
```