---
title: Star Wars Baby Names Sad Plot 2.0
author: Alex Michel
date: '2018-06-20'
slug: star-wars-baby-names-sad-plot-2-0
categories:
  - blog
tags:
  - baby names
  - R Markdown
  - sad plot
  - Star Wars
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

### 1) Clean Reproducible Code for Sad Plot 2.0

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(babynames)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(plotly)
library(RColorBrewer)

star_wars <- babynames %>%
  filter(name == "Leah" | name == "Leia" | name == "Padme") %>% 
  filter(year >= 1970) %>% 
  filter(sex == "F")

boy_names <- babynames %>% 
  filter(name == "Luke" | name == "Anakin" | name == "Han") %>% 
  filter(year >= 1970) %>% 
  filter(sex == "M")


plot <- ggplot(star_wars, aes(x = year, 
                           y = prop,  
                           group = name, sex,
                           color = name)) + 
  geom_line(size = 1) +
  geom_line(data=boy_names, size = 1) +
  geom_vline(xintercept = 1977, linetype="dashed", 
                color = "gray70", size=.8) +
  geom_vline(xintercept = 1999, linetype="dashed", 
                color = "gray70", size=.8) +
  geom_vline(xintercept = 2014, linetype="dashed", 
                color = "gray70", size=.8) +
  annotate("text", x = 1980.5, y = 0.006, label = "Star Wars") +
  annotate("text", x = 1993, y = 0.006, label = "Phantom Menace") +
  annotate("text", x = 2008.5, y = 0.006, label = "Force Awakens") +
  ggtitle("Does Star Wars Influence Baby Name Popularity?") +
  xlab("Year") + ylab("Proportion of Baby Names") +
  coord_cartesian(ylim = c(0, 0.006)) +
  scale_color_discrete(name="Baby Names",
                       breaks=c("Luke","Leah","Leia", "Anakin", "Han", "Padme")) +
    theme_minimal()

plot

```

### 2) What Type of Graph?

This is a a very simple line graph. 


### 3) Data Description

This graph was pretty much my first time graphing anything in ggplot2! 

The data came from the babynames package and the information for which Star Wars baby names to look for came from a Huffington Post article on the relative popularity of Star Wars inspired baby names. 

### 4) Intended Audience

Honestly there isn't much of an intended audience for this one (me?). I think it could be most useful as an accompanying illustration for an article on the rising popularity of Star Wars baby names.

### 5) What Am I Trying to Show?

1) Do Star Wars inspired baby names become more popular after a new movie comes out? 

2) Have Star Wars inspired baby names become "mainstream"? Do they have longterm staying power even years after a movie release?


### 6) How Do I Read This?

The main components of the graph are time and proportion of baby names. As the proportion increases the line ascends vertically. The use of time on the x-axis allows us to see whether the change in proportion is associated with particular points in time when particular Star Wars movies were released.

### 7) Presentation Tips

- Point out that the graph actually starts at 1970, 7 years before the first Star Wars was released. The reason for this is to give the viewer some context for the popularity of these names before the movies.

- Point out that even though the name Leah sounds like Leia the spelling for Leah is pretty common. It's also interesting that Leah seems to have experienced a bit of a spike in popularity around 2008 -- not sure why.


### 8) Variations and Alternatives

An alternative to this line graph could be a bar graph -- which would make it easier to see suddent changes associated with particular years.


### 9) How I Created It: Methods Section

This was pretty much the first time I've ever used ggplot2! I started off by copying Alison's example and filtering the data for some simple name variations (Leah vs Leia) on spelling. I filtered by sex and year (only including years after Star Wars came out).


#### Babynames Sad Graph 1.0

```{r}
star_wars <- babynames %>%
  filter(name == "Leah" | name == "Leia") %>% 
  filter(sex == "F") %>% 
  filter(year >= 1975)

plot <- ggplot(star_wars, aes(x = year, 
                           y = prop,  
                           group = name, 
                           color = name)) + 
  geom_line(size = 2) +
  geom_vline(xintercept = 1977, linetype="dashed", 
                color = "blue", size=1) +
  geom_vline(xintercept = 1999, linetype="dashed", 
                color = "blue", size=1) +
  geom_vline(xintercept = 2014, linetype="dashed", 
                color = "blue", size=1) +
  annotate("text", x = 1980, y = 0.01, label = "Star Wars") +
  annotate("text", x = 2004, y = 0.008, label = "Phantom Menace") +
  annotate("text", x = 2009.5, y = 0.005, label = "Force Awakens")

plot
```

This is like, a really sad plot. But, it's like basically my very first baby step. I think it is successful at communicating the intended information, but it's really ugly. Most of my changes to this plot focused on improving the appearance.

I started off version 2.0 by adding more information (more names) to help see if there was really a general trend in naming. Then, I changed the colors, the line size, arranged the annotation text to be more consistent.

The main title, axis titles, and legend all needed some tweaking as well. I could have really kept playing with this (changing it to a bar plot or tweaking the theme) but I also wanted to keep it similar enough to the original sad plot that it was still comparable.

