---
title: Star Wars Baby Names Sad Plot 2.0
author: Alex Michel
date: '2018-06-20'
slug: star-wars-baby-names-sad-plot-2-0
categories:
  - blog
tags:
  - baby names
  - R Markdown
  - sad plot
  - Star Wars
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(babynames)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(plotly)
library(RColorBrewer)
```

### Babynames Sad Graph

This graph was pretty much my first time graphing something in ggplot2! 

I would like to polish it up by:

1) Adding a few more names to it to make it a little more interesting (Luke, Han, Lando...). 
2) I'd like to change it to a density plot (I think?) to make it a little cooler looking. Because these names are probably a fairly small proportion of the total babynames, I'd like to do something to adjst the scale so it's easier to see. I'm thinking that a density plot of proportion of total Star Wars themed baby names, rather than just proportion of total American baby names, could be one way to do this.

```{r}
star_wars <- babynames %>%
  filter(name == "Leah" | name == "Leia") %>% 
  filter(sex == "F") %>% 
  filter(year >= 1975)

plot <- ggplot(star_wars, aes(x = year, 
                           y = prop,  
                           group = name, 
                           color = name)) + 
  geom_line(size = 2) +
  geom_vline(xintercept = 1977, linetype="dashed", 
                color = "blue", size=1) +
  geom_vline(xintercept = 1999, linetype="dashed", 
                color = "blue", size=1) +
  geom_vline(xintercept = 2014, linetype="dashed", 
                color = "blue", size=1) +
  annotate("text", x = 1980, y = 0.01, label = "Star Wars") +
  annotate("text", x = 2004, y = 0.008, label = "Phantom Menace") +
  annotate("text", x = 2009.5, y = 0.005, label = "Force Awakens")

plot
```

