---
title: Waco Home Prices Sad Plot 2.0
author: Alex Michel
date: '2018-06-20'
slug: waco-home-prices-sad-plot-2-0
categories:
  - blog
tags:
  - plot
  - R Markdown
  - sad plot
  - Waco
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

### 1) Clean Reproducible Code

```{r include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(plotly)
library(RColorBrewer)

waco <- txhousing %>% 
  filter(city == "Waco", year >= 2004) %>% 
  group_by(year) %>% 
  mutate(median_waco = median(median), 
         gt_median = median >= median_waco)

waco_homes <- ggplot(waco, aes(x = year, y = median, color = gt_median)) + 
    geom_point(alpha = 0.7) + labs(title = "Waco, TX median Sale Price Over Time", 
    subtitle = "Note, the show Fixer Upper debuted in 2013", caption = "Source: ggplot2 txhousing data") + 
    xlab("Year") + 
    ylab("Median sale price by month") + 
  scale_x_continuous(breaks = seq(2004, 2015, by = 2)) +
  scale_y_continuous(breaks = seq(90000, 175000, by = 10000)) +
  scale_color_discrete(name="Yearly Median Home Prices",
                      breaks=c("FALSE", "TRUE"),
                      labels=c("Below Median", "Above Median")) 
  

waco_homes +
  theme_minimal() 

```

### 2) What Type of Graph?

This is a scatterplot. 


### 3) Data Description

The data for this graph was pre-included in the R package. The data includes quantitative data on real estate prices for housing in Texas, including the city of Waco. The data has been pre-cleaned so that it is easy to work with right out of the box. The raw data includes 11 variables with 139 observations, with a .

### 4) Intended Audience

This could be useful for a homebuyer or a real estate agent, someone interested in whether a particular house is priced high or low for the overall market.

### 5) What Am I Trying to Show?

There are two main things that I wanted to show with this graph:

1) The general trend in Waco real estate -- which is clearly going up over time.

2) Whether particular houses are above or below the median valuation for Waco that year (i.e., is this a good deal or not?). Obviously there's a lot of other factors to consider, but it's an interesting way to see the market differences between "low value" real estate and "high value" real estate over time.

### 6) How Do I Read This?

This scatterplot graph uses a cartesian coordinate system where the year is along the x-axis and the median sale prices are along the y-axis. Each dot represents the median sale price for all houses sold during one month. This graph type is helpful at showing the relationship between house pricing and time -- it's very clear that home prices (even lower valued ones) are going up over time regardless of the economic recession from 2008.

### 7) Presentation Tips

It's important to note that the dots in the plot are monthly medians, so if I had more time to get to play with analyzing this I would want to know whether particular months are just bad for sales (maybe including some lines grouped by month?).


### 8) Variations and Alternatives

There are so many different variations for this type of graph. One that I see a lot is the use of bubbles that are sized according to the magnitude of values such that larger bubbles represent larger values.


### 9) How I Created It Methods Section

I initially made this graph as part of the color lab. I initially used continuous colors to help illustrate changes in price overall (higher prices were darker red, lowest prices were yellow). 

I started off with the dataset for all of Texas and decided to filter just for Waco after a 24-year old classmate asked me "What's Waco?". Also, in Cecile Richards' new book she talks about Waco as being famous for Chip and Johanna and HGTV, which is uh...not what the city of Waco brings to mind for me. 

Anyway, I started with a simple scatterplot using median house sale price and year variables. I noticed that there were several years in the dataset that didn't have the needed values, so I filtered for years after 2004. I played around with some of these values until I got a satisfactory looking scatterplot, then I added color using the ColorBrewer package.

```{r}
waco <- txhousing %>% 
  filter(city == "Waco", year >= 2004) %>% 
  group_by(year) 

waco_homes <- ggplot(waco, aes(x = year, y = median, color = median)) + 
    geom_point(alpha = 0.7) + labs(title = "Waco, TX median sale price over time", 
    subtitle = "Note, the show Fixer Upper debuted in 2013", caption = "Source: ggplot2 txhousing data") + 
    xlab("Year") + 
    ylab("Median sale price") + 
  scale_x_continuous(breaks = seq(2004, 2015, by = 2)) +
  scale_y_continuous(breaks = seq(90000, 175000, by = 10000)) 

waco_homes +
  scale_color_gradientn(colours = brewer.pal(n=7, name="YlOrRd")) +
  theme_minimal()
```


With help from Emile during our lab, I created a new indicator variable (a logical TRUE or FALSE) using mutate to show whether the home prices in a given year were above or below a threshold of the median home price for that year (which is the sad plot 2.0 at the top of the page). One of the things I like about the 2.0 version of the graph is that the viewer can compare across the full timescale as well as across individual years much more easily. The continuous color is helpful for a big picture impression (and it looks pretty) but it ultimately doesn't include as much information. 